---
title: "Human Activity Recognition"
author: "Shanmuga"
date: "20 July 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Human Activity Recognition

## Load raw data

```{r}
library(caret)
library(rattle)
library(e1071)
library(ggplot2)
library(C50)

trainData <- read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!",""))
testData <- read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!",""))
```

## Clean the raw data

Remove first column that is not required - running number
```{r}
cleanTrainData <- trainData[c(-1)]
```

Find NA values and exclude
```{r}
notNA <- sapply(cleanTrainData,function(i){sum(is.na(i))/length(i)})<0.9
cleanTrainData <- cleanTrainData[, notNA]
```

Remove near zero variance columns/variables
```{r}
nzv <- nearZeroVar(cleanTrainData, saveMetrics=TRUE)
cleanTrainData <- cleanTrainData[,nzv$nzv==FALSE]
```

Drop unnecessary columns from the testing dataset
```{r}
columnNames <- colnames(cleanTrainData[,-58]) # remove class column.
cleanTestData <- testData[columnNames]
```

Split the training set into a training and validation set.
```{r}
intrain<-createDataPartition(y=cleanTrainData$classe,p=0.7,list=FALSE)
training<-cleanTrainData[intrain,]
testing<-cleanTrainData[-intrain,]
```


### Cross Validation to select the best model performance

```{r}
fitControl <- trainControl(method = "cv", number = 10, repeats = 0,savePred=T, classProb=T)

modelHAR_1 <- train(classe~., data = training, method = "rpart", trControl= fitControl)
modelHAR_2 <- train(classe~., data = training, method = "rpart1SE", trControl= fitControl)
modelHAR_3 <- train(classe~., data = training, method = "C5.0Tree", trControl= fitControl)
```

Confusion martix
```{r}
activity_1 <- predict(modelHAR_1, newdata = testing[,-58])
conMatrix_1 <- confusionMatrix(activity_1, testing$classe)
conMatrix_1
```

```{r}
activity_2 <- predict(modelHAR_1, newdata = training[,-58])
conMatrix_2 <- confusionMatrix(activity_2, training$classe)
conMatrix_2
```

```{r}
activity_3 <- predict(modelHAR_3, newdata = training[,-58])
conMatrix_3 <- confusionMatrix(activity_3, training$classe)
conMatrix_3
```
As the "C5.0Tree" has the highest accuracy (0.9972) when predicting the test data. we select this model to build our final model.


### Build final model
Final model is build with the "C5.0Tree" algorithm using the entire training data set.

```{r}
fitControl <- trainControl(method = "cv", number = 10, repeats = 0,savePred=T, classProb=T)
modelHAR_F <- train(classe~., data = cleanTrainData, method = "C5.0Tree", trControl= fitControl)
```


### Perform prediction with final model
```{r}
act_prediction <- predict(modelHAR_F, newdata = cleanTestData)

```

The predistions obtained are:
```{r}
act_prediction
```







